\tongjisetup{
  %******************************
  % 注意：
  %   1. 配置里面不要出现空行
  %   2. 不需要的配置信息可以删除
  %******************************
  %
  %=====
  % 秘级
  %=====
  secretlevel={保密},
  secretyear={2},
  %
  %=========
  % 中文信息
  %=========
  % 题目过长可以换行。
  ctitle={面向无人机自主堆体体积测量的视觉定位及重建方法研究},
  cheadingtitle={面向无人机自主堆体体积测量的视觉定位及重建方法研究},    %用于页眉的标题，不要换行
  cauthor={何士波},  
  studentnumber={1732940},
  cmajorfirst={工学},
  cmajorsecond={控制工程},
  cdepartment={电子与信息工程学院},
  csupervisor={岳继光~教授}, 
  % 如果没有副指导老师或者校外指导老师，把{}中内容留空即可，或者直接注释掉。
  cassosupervisor={董延超~副教授}, % 副指导老师
  % 日期自动使用当前时间，若需指定按如下方式修改：
  % cdate={超新星纪元},
  % 没有基金的话就注释掉吧。
  %cfunds={（国家杰出青年基金 (No.123456789) 支持）},
  %
  %=========
  % 英文信息
  %=========
  etitle={Visual localization and reconstruction method for autonomous bulk measurement of unmanned aerial vehicle}, 
  eauthor={He ShiBo},
  emajorfirst={Engineering},
  emajorsecond={Control Engineering},
  edepartment={College of Electronics and Information Engineering},
  %efunds={(Supported by the Natural Science Foundation of China for\\ Distinguished Young Scholars, Grant No.123456789)},    
  esupervisor={Prof. Yue Jiguang},
  eassosupervisor={V.P. Dong Yanchao}
  }

% 定义中英文摘要和关键字
\begin{cabstract}  
  在现代工业中，随着工业生产水平的提高，为了加强对堆体物料库存量的监控，针对堆体积测量的需求也逐步提升。传统测量方案一般都是选用激光雷达，测重地磅等复杂仪器来完成测量工作，这些仪器可以一定程度的完成测量工作并获取测量真实值。但依旧存在很多问题限制以上测量技术在行业内的广泛应用，例如测量仪器成本过高，安装复杂，单一场景内的仪器难以复用；测量过程需要人力参与，难以实现全自动化的测量流程；测量精度年久失衡，难以保证测量精度。针对以上情况，本文提出了一种在封闭电厂环境下，通过无人机的自主定位和飞行，对煤堆体进行图像采集工作，随后根据采集的图像进行三维重建得到高精度堆体点云模型，对点云模型估计体积值以获取堆体体积的完整测量流程。本文所提出方案完全基于计算机视觉实现，成本可控，且能在各种场景中复用。
  
  针对无人机的自主定位和飞行方法研究。本文提出在封闭无GPS的环境中，使用纯视觉的方式来完成无人机的自主定位和飞行工作。通过在场景中放置二维码，来解决视觉SLAM无法确定尺度的问题，并且结合SLAM坐标系和二维码坐标系对无人机提供真实世界坐标系下的位姿信息。
  
  针对连续图像的高精度堆体三维重建方法研究。本文提出一种结合SLAM结果的堆体三维重建方法，以有序图像作为SLAM输入，关键帧数据集作为输出提供给三维重建系统以获取高精度点云。解决传统三维重建图像匹配耗时，点云精度低的问题。
  
  针对堆体点云进行体积估计方法研究。本文提出一种基于计算机视觉方案的堆体体积估计方法，依次对生成的点云进行滤波获取高精度点云，估计三维点云的实际尺度，确定点云模型水平面等工作，最后根据纯点云信息估算场景中堆体的体积大小。
  
  基于真实环境下，对堆体体积测量进行分模块测试。实验结果表明，无人机的定位误差在高度方向可以控制在0.2m以内，整体定位精度在5$\%$以内，该精度完全可以提供给无人机循迹使用；改进后的三维重建模块则能加快匹配流程，获取较高精度的堆体三维点云，以满足体积测量的需求；最后的体积测量模块，可以估计出准确的尺度大小和水平面解析方程，通过对点云体积的求解获得堆体体积，测量误差可以控制在2$\%$以内。
\end{cabstract}

\ckeywords{立体视觉,SLAM,无人机定位, 三维重建,堆体体积测量}

\begin{eabstract}
  In modern industry, with the improvement of industrial production level, in order to strengthen the monitoring of heap material inventory, the demand for heap volume measurement is also gradually increasing.Traditional measurement schemes usually use laser radar, weighometer and other complex instruments to complete the measurement work, these instruments can complete the measurement work to a certain extent and obtain the real value of the measurement.However, there are still many problems that limit the wide application of the above measurement technology in the industry.The measurement process needs human participation, so it is difficult to realize the fully automated measurement process;The measurement accuracy is out of balance for years, so it is difficult to guarantee the measurement accuracy.In view of the above situation, this paper puts forward a closed environment of power plant, through independent positioning and flight of the unmanned aerial vehicle (uav), the pile body image acquisition work, then according to the collected image three dimensional reconstruction of pile body point cloud model has high precision, the point cloud model to estimate the volume value for the pile body volume of the complete measuring process.The scheme proposed in this paper is completely based on computer vision, with controllable cost, and can be reused in various scenarios.
  
  Research on autonomous positioning and flight method of uav.This paper proposes to use pure vision to complete the autonomous positioning and flight of uav in closed environment without GPS.The two-dimensional code is placed in the scene to solve the problem that visual SLAM cannot determine the scale, and the combination of the SLAM coordinate system and the two-dimensional code coordinate system provides the positioning information of the uav in the real world coordinate system.
  
  Research on 3d reconstruction of continuous image with high precision.In this paper, a method of three-dimensional reconstruction of the heap combined with SLAM results is proposed, in which the ordered image is taken as the input of SLAM, and the key frame data set is provided to the three-dimensional reconstruction system as the output to obtain the high-precision point cloud.To solve the problem of time-consuming matching of traditional 3d reconstruction image and low accuracy of point cloud.
  
  The volume estimation method of the point cloud is studied.In this paper, a method of heap volume estimation based on computer vision scheme is proposed, which successively filters the generated point cloud to obtain the high-precision point cloud, estimates the actual scale of the three-dimensional point cloud, determines the horizontal plane of the point cloud model, and finally estimates the volume of the heap in the scene based on the pure point cloud information.
  
  Based on the real environment, the volume measurement of the reactor is tested by module.The experimental results show that the positioning error of uav can be controlled within 0.2m in the altitude direction, and the overall positioning accuracy can be controlled within 5$\%$, which can be fully used for tracking of uav.The improved 3d reconstruction module can speed up the matching process and obtain 3d point cloud of the heap with high accuracy to meet the demand of volume measurement.Finally, the volume measurement module can estimate the accurate scale size and the analytic equation of the horizontal plane. The volume of the heap can be obtained by solving the volume of the point cloud, and the measurement error can be controlled within 2$\%$.
\end{eabstract}

\ekeywords{keywords}